---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tcopula

<!-- badges: start -->
<!-- badges: end -->

`tcopula` implements the algorithm described in the paper ["Estimation of Structured t-Copulas"](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=1126401), from 
Attilio Meucci.

## Installation

You can install the development version of `tcopula` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Reckziegel/tcopula")
```

## Case Study

This section replicates the original paper results. 

Consider constant maturities swaps of 7 different issue dates (`1y`, `2y`, `5y`, `7y`, `10y`, `15y` and `30y`), ranging from `1995-05-01` to `2005-04-21`, for a total of 2500 daily realizations. 

```{r}
library(tcopula)

data("DB_SwapParRates")

Rates <- DB_SwapParRates

head(Rates, 2)

tail(Rates, 2)
```

To model this data, first recognize that, as point out in [Risk and Asset Allocation](https://www.springer.com/gp/book/9783540222132), the daily changes in interest-rates can be considered ~ approximately ~ "invariant" (stationary). For 
that reason, the estimation process is usually carried out by taking the 
interest-rates first differences:

```{r}
X <- Rates[2:nrow(Rates), ] - Rates[1:nrow(Rates) - 1, ]
```

On top of that, it's known that the the first 3 principal components explain most 
of the yield-curve variations (see Litterman and Scheinkman paper [here](https://www.math.nyu.edu/faculty/avellane/Litterman1991.pdf)). 

We use this knowledge to shrink the dimension of the dataset and avoid noisy signals
from low eigenvalues that are not well defined. By setting `K = 3`, the main sources
of relevant information (aka level, steepness and curvature) are taken into account, while the remaining factors are assumed to follow and isotropic structure: 

```{r}
K <- 3
tol <- 10 ^ (-10)

fit <- StrucTMLE(X = X, K = K, Tolerance = tol)
fit 
```

As the output shows, the commovements in interest-rates variations are 
quite high. The optimum degree of freedom is $v=50$, which may indicate that 
fat-tails are not an issue for this particular dataset. 

Finally, `figure 1` can be replicated with an additional help from 
the [ggplot2](https://ggplot2.tidyverse.org/) package.

```{r}
library(ggplot2)

df <- data.frame(
  factors     = as.factor(1:7), 
  eigenvalues = eigen(fit$C)$values
) 

ggplot(data = df, aes(x = factors, y = eigenvalues, fill = factors)) + 
  geom_col() + 
  scale_fill_viridis_d(option = "E")

```


